Package ‘cdata’
October 21, 2018
Type Package
Title Fluid Data Transformations
Version 1.0.3
Date 2018-10-20
URL https://github.com/WinVector/cdata/,
https://winvector.github.io/cdata/
Maintainer John Mount <jmount@win-vector.com>
BugReports https://github.com/WinVector/cdata/issues

#### Description 
 Supplies higher-order fluid data transform operators that include pivot and anti-
pivot as special cases.
The methodology is describe in 'Zumel', 2018, ``Fluid data reshap-
ing with 'cdata''', <http://winvector.github.io/FluidData/FluidDataReshapingWithCdata.html> , doi:10.5281/zenodo.117329
Based on the 'DBI' database interface.
License GPL-3
Encoding UTF-8
LazyData true
RoxygenNote 6.1.0
Depends R (>= 3.4.0)
Imports wrapr (>= 1.6.3), rquery (>= 1.1.0), stats
Suggests DBI, RSQLite, testthat, knitr, rmarkdown, data.table (>=
1.11.4), rqdatatable (>= 1.1.1)
VignetteBuilder knitr
ByteCompile true
NeedsCompilation no
Author John Mount [aut, cre],
Nina Zumel [aut],
Win-Vector LLC [cph]
Repository CRAN
Date/Publication 2018-10-21 07:40:03 UTC
12
blocks_to_rowrecs
R topics documented:
blocks_to_rowrecs . . . . .
blocks_to_rowrecs_q . . . .
build_pivot_control . . . . .
build_pivot_control_q . . . .
build_unpivot_control . . . .
cdata . . . . . . . . . . . . .
checkColsFormUniqueKeys
map_fields . . . . . . . . . .
map_fields_q . . . . . . . .
pivot_to_rowrecs . . . . . .
qlook . . . . . . . . . . . .
rowrecs_to_blocks . . . . .
rowrecs_to_blocks_q . . . .
unpivot_to_blocks . . . . . 




Index
blocks_to_rowrecs



Map sets rows to columns (takes a data.frame).

#### Description 

Transform data facts from rows into additional columns using controlTable.

#### Usage 
 <pre><code>
blocks_to_rowrecs(tallTable, keyColumns, controlTable, ...,
columnsToCopy = NULL, checkNames = TRUE, checkKeys = TRUE,
strict = FALSE, use_data_table = FALSE)
## Default S3 method:
blocks_to_rowrecs(tallTable, keyColumns, controlTable,
..., columnsToCopy = NULL, checkNames = TRUE, checkKeys = TRUE,
strict = FALSE, use_data_table = TRUE)
## S3 method for class 'relop'
blocks_to_rowrecs(tallTable, keyColumns, controlTable, ...,
columnsToCopy = NULL, checkNames = TRUE, strict = FALSE,
use_data_table = TRUE,
tmp_name_source = wrapr::mk_tmp_name_source("bltrr"),
temporary = TRUE)blocks_to_rowrecs
3
</code></pre>
#### Arguments 

tallTable data.frame containing data to be mapped (in-memory data.frame).
keyColumns character vector of column defining row groups
controlTable table specifying mapping (local data frame)
... force later arguments to be by name.
columnsToCopy character, extra columns to copy.
checkNames logical, if TRUE check names.
checkKeys logical, if TRUE check keyColumns uniquely identify blocks (required).
strict logical, if TRUE check control table name forms
use_data_table logical if TRUE try to use data.table for the pivots.
tmp_name_source
a tempNameGenerator from cdata::mk_tmp_name_source()
temporary
logical, if TRUE make result temporary.

#### Details 

This is using the theory of "fluid data"n (https://github.com/WinVector/cdata), which in-
cludes the principle that each data cell has coordinates independent of the storage 
#### Details 
 and stor-
age detail dependent coordinates (usually row-id, column-id, and group-id) can be re-derived at will
(the other principle is that there may not be "one true preferred data shape" and many re-shapings
of data may be needed to match data to different algorithms and methods).
The controlTable defines the names of each data element in the two notations: the notation of the
tall table (which is row oriented) and the notation of the wide table (which is column oriented).
controlTable[ , 1] (the group label) cross colnames(controlTable) (the column labels) are names of
data cells in the long form. controlTable[ , 2:ncol(controlTable)] (column labels) are names of data
cells in the wide form. To get behavior similar to tidyr::gather/spread one builds the control table
by running an appropriate query over the data.

Some discussion and examples  can be found here: https://winvector.github.io/FluidData/
FluidData.html and here https://github.com/WinVector/cdata.
Value
wide table built by mapping key-grouped tallTable rows to one row per group
See Also
build_pivot_control, rowrecs_to_blocks_q


#### Examples 
```{r}
# pivot example
d <- data.frame(meas = c('AUC', 'R2'),
val = c(0.6, 0.2))
cT <- build_pivot_control(d,4
blocks_to_rowrecs_q
columnToTakeKeysFrom= 'meas',
columnToTakeValuesFrom= 'val')
blocks_to_rowrecs(d,
keyColumns = NULL,
controlTable = cT)
d <- data.frame(meas = c('AUC', 'R2'),
val = c(0.6, 0.2))
cT <- build_pivot_control(
d,
columnToTakeKeysFrom= 'meas',
columnToTakeValuesFrom= 'val')
ops <- rquery::local_td(d) %.>%
blocks_to_rowrecs(.,
keyColumns = NULL,
controlTable = cT)
cat(format(ops))
if(requireNamespace("rqdatatable", quietly = TRUE)) {
library("rqdatatable")
d %.>%
ops %.>%
print(.)
}
if(requireNamespace("RSQLite", quietly = TRUE)) {
db <- DBI::dbConnect(RSQLite::SQLite(), ":memory:")
DBI::dbWriteTable(db,
'd',
d,
overwrite = TRUE,
temporary = TRUE)
db %.>%
ops %.>%
print(.)
DBI::dbDisconnect(db)
}
```


## ``blocks_to_rowrecs_q``: Map sets rows to columns (query based, take name of table).

#### Description 

Transform data facts from rows into additional columns using SQL and controlTable.

#### Usage 
 <pre><code>
blocks_to_rowrecs_q(tallTable, keyColumns, controlTable, my_db, ...,
columnsToCopy = NULL,
tempNameGenerator = mk_tmp_name_source("mvtcq"), strict = FALSE,
checkNames = TRUE, showQuery = FALSE, defaultValue = NULL,
dropDups = FALSE, temporary = FALSE, resultName = NULL)
</code></pre>
#### Arguments 

tallTable name of table containing data to be mapped (db/Spark data)
keyColumns character list of column defining row groups
controlTable table specifying mapping (local data frame)
my_db db handle
... force later arguments 
 to be by name.
* ``columnsToCopy``: character list of column names to copy
* tempNameGenerator: a tempNameGenerator from cdata::mk_tmp_name_source()
strict logical, if TRUE check control table name forms
checkNames logical, if TRUE check names
showQuery if TRUE print query
defaultValue if not NULL literal to use for non-match values.
dropDups logical if TRUE suppress duplicate columns (duplicate determined by name, not
content).
temporary logical, if TRUE make result temporary.
resultName character, name for result table.

#### Details 

This is using the theory of "fluid data"n (https://github.com/WinVector/cdata), which in-
cludes the principle that each data cell has coordinates independent of the storage 
#### Details 
 and stor-
age detail dependent coordinates (usually row-id, column-id, and group-id) can be re-derived at will
(the other principle is that there may not be "one true preferred data shape" and many re-shapings
of data may be needed to match data to different algorithms and methods).
The controlTable defines the names of each data element in the two notations: the notation of the
tall table (which is row oriented) and the notation of the wide table (which is column oriented).
controlTable[ , 1] (the group label) cross colnames(controlTable) (the column labels) are names of
data cells in the long form. controlTable[ , 2:ncol(controlTable)] (column labels) are names of data
cells in the wide form. To get behavior similar to tidyr::gather/spread one builds the control table
by running an appropriate query over the data.



#### Examples 
```{r}
if (requireNamespace("DBI", quietly = TRUE) &&
requireNamespace("RSQLite", quietly = TRUE)) {
my_db <- DBI::dbConnect(RSQLite::SQLite(), ":memory:")
# pivot example
d <- data.frame(meas = c('AUC', 'R2'), val = c(0.6, 0.2))
DBI::dbWriteTable(my_db,
'd',
d,
temporary = TRUE)
cT <- build_pivot_control_q('d',
columnToTakeKeysFrom= 'meas',
columnToTakeValuesFrom= 'val',
my_db = my_db)
tab <- blocks_to_rowrecs_q('d',
keyColumns = NULL,
controlTable = cT,
my_db = my_db)
qlook(my_db, tab)
DBI::dbDisconnect(my_db)
}
```

Some discussion and examples   can be found here: https://winvector.github.io/FluidData/
FluidData.html and here https://github.com/WinVector/cdata.
Value
wide table built by mapping key-grouped tallTable rows to one row per group6
build_pivot_control
See Also
rowrecs_to_blocks_q, build_pivot_control_q, blocks_to_rowrecs
