---
title: "Introduction to ggplot2"
author: "Daniel McInerney, Kevin O'Brien"
date: "18 February 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Getting Started with ggplot2

This R markdown file is an introduction to data analysis and visualisation using tidyverse, specifically dplyr and ggplot2:

The exercise uses a subset of the Sub-compartment database, which has been exported as a comma separated file (CSV). 

### Data Import
Let's start by reading in the data using `read.csv`. Note that you can specify if the file has a header and what the field separator is.
```{r}
scdb <- read.csv('../DATA/scdb_subset.csv', head=TRUE, sep=',')
summary(scdb)
```


### First Graphic
Load the ggplot2 library

```{r}
library(ggplot2)
```

Let's start by creating a bar chart of counts of species within the database.

```{r}
ggplot(data=scdb) + geom_bar(mapping = aes(x=spp1))
```


Let's make it more readable by limiting the chart to **spruce** only:

```{r}
ggplot(data=dplyr::filter(scdb, spp1=='SS'|spp1=='NS')) + geom_bar(mapping = aes(x=spp1))
```


Let's extend this idea by creating a bar chart using two variables: Yield Class and Species.

```{r}
ggplot(data=scdb, aes(yc1, fill=spp1)) +  geom_histogram()
```

This is a start, but a bit overloaded.

**Exercise**:  From what you learnt earlier, re-create the graphic, but only create a histogram using SS. (hint: you can use the filter function).



```{r, eval=TRUE}
ggplot(data=dplyr::filter(scdb, spp1=='SS'), aes(yc1, fill=spp1)) + geom_histogram()
```


```{r, eval=TRUE}
ggplot(data=scdb, aes(x=spp1, y=gis_area)) + geom_boxplot()
```

