Basic Plots 2
========================================================

```{r}
library(ggplot2)
```

### Plotting a Function
- To plot a function, use `stat_function()`. 
- It's also necessary to give `ggplot()` a dummy data frame so that
it will get the proper x range. 
- In this example we'll use `dnorm()`, which gives the density
of the normal distribution.

```{r fig.width=10, fig.height=4}
# The data frame is only used for setting the range
p <- ggplot(data.frame(x=c(-3,3)), aes(x=x))
p + stat_function(fun = dnorm)
```
#### User Defined Functions

```{r fig.width=10, fig.height=5}
#Define a Sigmoid Function
myfun <- function(xvar) {
1/(1 + exp(-xvar + 10))
}
ggplot(data.frame(x=c(0, 20)), aes(x=x)) + stat_function(fun=myfun)
```
<hr>

### Shading a Subregion Under a Function Curve
Define a new wrapper function around your curve function, and replace out-of-range values with NA()
```{r fig.width=10, fig.height=4}
# Return dnorm(x) for 0 < x < 2, and NA for all other x
dnorm_limit <- function(x) {
y <- dnorm(x)
y[x < 0 | x > 2] <- NA
return(y)
}
# ggplot() with dummy data
p <- ggplot(data.frame(x=c(-3, 3)), aes(x=x)) + stat_function(fun=dnorm)
#Same plot as before.
```
```{r fig.width=10, fig.height=4}
p + stat_function(fun=dnorm_limit, geom="area", fill="blue", alpha=0.2)
```

### Creating a Graph of an Empirical Cumulative Distribution Function

```{r fig.width=10, fig.height=4}

library(gcookbook) # For the data set
# ecdf of heightIn
ggplot(heightweight, aes(x=heightIn)) + stat_ecdf()
# ecdf of ageYear
ggplot(heightweight, aes(x=ageYear)) + stat_ecdf()
```

### Network Graphs

```{r fig.width=10, fig.height=4}
# May need to install first, with install.packages("igraph")
library(igraph)
# Specify edges for a directed graph
gd <- graph(c(1,2, 2,3, 2,4, 1,4, 5,6, 4,3, 5,3, 2,6))
plot(gd)
# For an undirected graph
gu <- graph(c(1,2, 2,3, 2,4, 1,4, 5,2, 5,6, 3,6), directed=FALSE)
# No labels
plot(gu, vertex.label=NA)
```

