`geom_histograms`
========================================================

```{r fig.width=7, fig.height=6}
library(ggplot2)
set.seed(5689)
movies <- movies[sample(nrow(movies), 1000), ]
# Simple examples
qplot(rating, data=movies, geom="histogram")
qplot(rating, data=movies, weight=votes, geom="histogram")
qplot(rating, data=movies, weight=votes, geom="histogram", binwidth=1)
qplot(rating, data=movies, weight=votes, geom="histogram", binwidth=0.1)
```
```{r fig.width=7, fig.height=6}
# More complex
m <- ggplot(movies, aes(x=rating))
m + geom_histogram()
m + geom_histogram(aes(y = ..density..)) + geom_density()
```
```{r fig.width=7, fig.height=6}
m + geom_histogram(binwidth = 1)
m + geom_histogram(binwidth = 0.5)
m + geom_histogram(binwidth = 0.1)
```
```{r fig.width=7, fig.height=6}
# Add aesthetic mappings
m + geom_histogram(aes(weight = votes))
m + geom_histogram(aes(y = ..count..))
m + geom_histogram(aes(fill = ..count..))
```
```{r fig.width=7, fig.height=6}
# Change scales
m + geom_histogram(aes(fill = ..count..)) +
  scale_fill_gradient("Count", low = "green", high = "red")
```
Often we don't want the height of the bar to represent the
count of observations, but the sum of some other variable.
For example, the following plot shows the number of movies
in each rating.
```{r fig.width=7, fig.height=6}

qplot(rating, data=movies, geom="bar", binwidth = 0.1)

```
If, however, we want to see the number of votes cast in each
category, we need to weight by the votes variable
```{r fig.width=7, fig.height=6}
qplot(rating, data=movies, geom="bar", binwidth = 0.1,
  weight=votes, ylab = "votes")
```
```{r fig.width=7, fig.height=6}
m <- ggplot(movies, aes(x = votes))

```
### Transformed Scales
For transformed scales, binwidth applies to the transformed data.
The bins have constant width on the transformed scale.
```{r fig.width=7, fig.height=6}
m + geom_histogram() + scale_x_log10()
m + geom_histogram(binwidth = 1) + scale_x_log10()
m + geom_histogram() + scale_x_sqrt()
m + geom_histogram(binwidth = 10) + scale_x_sqrt()
```
```{r fig.width=7, fig.height=6}
# For transformed coordinate systems, the binwidth applies to the
# raw data.  The bins have constant width on the original scale.

# Using log scales does not work here, because the first
# bar is anchored at zero, and so when transformed becomes negative
# infinity.  This is not a problem when transforming the scales, because
# no observations have 0 ratings.
```
```{r fig.width=7, fig.height=6}
m + geom_histogram(origin = 0) + coord_trans(x = "log10")
# Use origin = 0, to make sure we don't take sqrt of negative values
m + geom_histogram(origin = 0) + coord_trans(x = "sqrt")
m + geom_histogram(origin = 0, binwidth = 1000) + coord_trans(x = "sqrt")
```
### Transforming the y axis.
You can also transform the y axis.  Remember that the base of the bars
has value 0, so log transformations are not appropriate
```{r fig.width=7, fig.height=6}

m <- ggplot(movies, aes(x = rating))
m + geom_histogram(binwidth = 0.5) + scale_y_sqrt()
m + geom_histogram(binwidth = 0.5) + scale_y_reverse()
```
### Setting aesthetics to fixed value
```{r fig.width=7, fig.height=6}

m + geom_histogram(colour = "darkgreen", fill = "white", binwidth = 0.5)
```
```{r fig.width=7, fig.height=6}
# Use facets
m <- m + geom_histogram(binwidth = 0.5)
m + facet_grid(Action ~ Comedy)
```
```{r fig.width=7, fig.height=6}
# Often more useful to use density on the y axis when facetting
m <- m + aes(y = ..density..)
m + facet_grid(Action ~ Comedy)
m + facet_wrap(~ mpaa)
```
```{r fig.width=7, fig.height=6}
# Multiple histograms on the same graph
# see ?position, ?position_fill, etc for more details.
set.seed(6298)
diamonds_small <- diamonds[sample(nrow(diamonds), 1000), ]
ggplot(diamonds_small, aes(x=price)) + geom_bar()
hist_cut <- ggplot(diamonds_small, aes(x=price, fill=cut))
hist_cut + geom_bar() # defaults to stacking
hist_cut + geom_bar(position="fill")
hist_cut + geom_bar(position="dodge")
```
```{r fig.width=7, fig.height=6}
# This is easy in ggplot2, but not visually effective.  It's better
# to use a frequency polygon or density plot.  Like this:
ggplot(diamonds_small, aes(price, ..density.., colour = cut)) +
  geom_freqpoly(binwidth = 1000)
# Or this:
ggplot(diamonds_small, aes(price, colour = cut)) +
  geom_density()
# Or if you want to be fancy, maybe even this:
ggplot(diamonds_small, aes(price, fill = cut)) +
  geom_density(alpha = 0.2)
# Which looks better when the distributions are more distinct
ggplot(diamonds_small, aes(depth, fill = cut)) +
  geom_density(alpha = 0.2) + xlim(55, 70)

```

