Scatterplots
========================================================

### 1. Preliminaries
- load **ggplot2**
- Inspect the ***iris*** data set
```{r}
library(ggplot2)
dim(iris)
names(iris)
class(iris)
table(iris$Species)
```

### 2. Scatterplots
- Basic scatterplot - Use `geom_point()`, and map one variable to x and one to y.
```{r fig.width=7, fig.height=6}
#plot(iris$Sepal.Length,iris$Petal.Width)
ggplot(iris, aes(x=Sepal.Length, y=Petal.Width)) + geom_point()
```
#### 2.1 Changing Plot Character
- Change the plot character to triangle (number 17).
- Change the colour to red.
- Increase size to 3. (The default value of `size` is 2, whereas default for `cex` is 1.)
```{r fig.width=7, fig.height=6}
#plot(iris$Sepal.Length,iris$Petal.Width, pch=17,col="red", cex=1.5)
ggplot(iris, aes(x=Sepal.Length, y=Petal.Width)) + geom_point(shape=17,col="red",size=3)
```

### 3. Grouping Data Points by a Variable Using Shape or Colour.
We can group points on the variable `Species` by mapping specied to one of the aesthetics 
- colour
- shape 
```{r fig.width=7, fig.height=6}
#ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width)) 
ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width, colour= Species)) + geom_point(size=3)
ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width, shape= Species)) + 
  geom_point(size=3)

```

Both of these approaches can be used in tandem. Lets use a data set with multiple categorical variables (***diamonds*** and ***tips*** data sets in the **ggplot2** and ***reshape2*** package respectively). The outcome is a bit hard to read.

```{r fig.width=7, fig.height=6}
ggplot(diamonds, aes(x=price, y=log(z), colour= clarity)) + geom_point(size=3)

ggplot(diamonds, aes(x=price, y=log(z), colour= clarity,shape=cut)) + geom_point(size=3)

```

Easier to look at.

```{r fig.width=7, fig.height=6}
library(reshape2)
ggplot(tips, aes(x=total_bill, y=tip, colour= day, shape = smoker)) + geom_point(size=3)


```


