Summarized Data Distributions
========================================================

Chapter 6 of ***R Graphics cookbook*** by Winston Chang

### 1. Preliminaries
- Load packages **ggplot2** and **MASS**
- Prepare Data Sets **faithful**(base) and **birthwt**(MASS)

```{r}
# Preliminaries
library(ggplot2)
library(MASS) # Birth Weight data set

# Make a copy of the data
birthwt1 <- birthwt
# Convert smoke to a factor
birthwt1$smoke <- factor(birthwt1$smoke)
```

#### 1.1 Old Faithful Data Set
Description: Waiting time between eruptions and the duration of the eruption for the Old Faithful geyser in Yellowstone National Park, Wyoming, USA.

- `eruptions` (numeric)	Eruption time in mins
- `waiting` (numeric)  Waiting time to next eruption (in mins)

```{r}
dim(faithful)
head(faithful)
```

#### 1.2 Birthweight Data Set
```{r}
dim(birthwt1)
head(birthwt1)
```
<hr>
### 2. Making a Basic Histogram
***Example 6.1***
```{r fig.width=7, fig.height=6}
ggplot(faithful, aes(x=waiting)) + geom_histogram()
```


#### 2.1 Adjusting binwidth
We will focus on an overview of the geoms in general, but this warning message indicated that these can take additional arguments. Let's change the binwidth to 3.
```{r fig.width=7, fig.height=6}
ggplot(faithful, aes(x=waiting)) + geom_histogram(binwidth=3)
# No Warning message
```
<hr>

### 3. Making a Density Curve
***Example 6.3***
- use of `geom_density()`
```{r fig.width=7, fig.height=6}
ggplot(faithful, aes(x=waiting)) + geom_density()
```
<hr>

```{r fig.width=7, fig.height=6}
ggplot(faithful, aes(x=waiting)) +
geom_line(stat="density", adjust=.25, colour="red") +
geom_line(stat="density") +
geom_line(stat="density", adjust=2, colour="blue")
```
<hr>
### 4. Making Multiple Density Curves from Grouped Data
***Example 6.5***
```{r fig.width=7, fig.height=6}

# Map smoke to colour
ggplot(birthwt1, aes(x=bwt, colour=smoke)) + geom_density()
# Map smoke to fill and make the fill semitransparent by setting alpha
ggplot(birthwt1, aes(x=bwt, fill=smoke)) + geom_density(alpha=.3)
```

<hr>
### 5. Making a Frequency Polygon
***Example 6.6***

- use of `geom_freqpoly()`

```{r fig.width=7, fig.height=6}
ggplot(faithful, aes(x=waiting)) + geom_freqpoly()
```
