Basic Plots
========================================================

### 1. Making a Basic Bar Graph

- X variable : categorical variable (Species variable in iris)
- Y variable : Some numeric variable 
- add `geom_bar`
```{r fig.width=7, fig.height=6}
library(ggplot2)
ggplot(iris, aes(x=Species, y=Sepal.Width)) + geom_bar(stat="identity")
```

#### 1.1 Playing around with different geoms.

```{r fig.width=7, fig.height=6}
# They may or may not be useful.

ggplot(iris, aes(x=Species, y=Sepal.Width)) + geom_point()

```

### 2. Fill
- By default, bar graphs use a very dark grey for the bars. To use a color fill, use `fill`.

- For the following example, we use a red fill and a black outline:

```{r fig.width=7, fig.height=3}

ggplot(iris, aes(x=Species, y=Sepal.Width)) + geom_bar(stat="identity", fill="red")

```
- Also, by default, there is no outline around the fill. To add an outline, use `colour` (Doesn't work all the time). 
```{r fig.width=7, fig.height=3}

ggplot(iris, aes(x=Species, y=Sepal.Width)) + geom_bar(stat="identity", colour="black")

```
