<style>
  p {     font-size:16px;
  }
</style>
qplot
========================================================

***qplot*** (short for quick plot) makes it easy to
produce complex plots, often requiring several lines of code using other plotting systems, in one line. 

***qplot()*** can do this because it's based on the grammar of graphics, which allows you to create a simple, yet expressive, description
of the plot.

***qplot()*** has been designed to be very similar to plot, which should make
it easy if you're already familiar with plotting in R. Remember, during an R
session you can get a summary of all the arguments to qplot with R help,
`?qplot`.

<hr>
### 1. Preliminaries
- Install package
- Inspect **mtcars** data set
```{r}
library(ggplot2)
# continuous scale vs. discrete scale
head(mtcars)
names(mtcars)
table(mtcars$cyl)
levels(mtcars$cyl)
```

```{r fig.width=7, fig.height=4}
qplot(wt, mpg, data=mtcars)

```
### Colour, size, shape and other aesthetic attributes

The first big difference when using qplot instead of plot comes when you
want to assign colours-or sizes or shapes-to the points on your plot. 

***qplot()*** makes it easy to include additional data on the plot.

#### Main Aesthetic Attributes
1. colour
2. size
3. shape

```{r fig.width=7, fig.height=4}
qplot(wt, mpg, data=mtcars, colour=cyl)

```

```{r fig.width=7, fig.height=4}
# Specifying cyl as a factor
qplot(wt, mpg, data=mtcars, colour=factor(cyl))

```
### Use Different "aesthetic"" mappings size and shape

```{r fig.width=7, fig.height=4}
qplot(wt, mpg, data=mtcars, shape=factor(cyl))
qplot(wt, mpg, data=mtcars, size=qsec)
```

### Combine mappings
```{r fig.width=7, fig.height=4}

# (hint: hollow points, geom-concept, legend combination)
qplot(wt, mpg, data=mtcars, size=qsec, color=factor(carb))
```
### Plot Characters
```{r fig.width=7, fig.height=4}
# Default pch (pch = 1) and pch = 17
qplot(wt, mpg, data=mtcars, size=qsec, color=factor(carb), shape=I(1))
qplot(wt, mpg, data=mtcars, size=qsec, color=factor(carb), shape=I(17))
```
Remark : `I()` inhibits interpretation/conversion of objects

### Scatterplot

```{r fig.width=7, fig.height=4}
qp1 <- qplot(wt, mpg, data=mtcars, size=qsec, shape=factor(cyl))
class(qp1)
qp1
```

Explicitly use `geom_point` - the reason will become clear shortly.
```{r fig.width=7, fig.height=4}

qplot(wt, mpg, data=mtcars, size=qsec, shape=factor(cyl), geom="point")
qplot(wt, mpg, data=mtcars, size=factor(cyl), geom="point")
```

### Enhancing plots

```{r fig.width=7, fig.height=4}
# bar-plot
qplot(factor(cyl), data=mtcars, geom="bar")
# flip plot by 90°
qplot(factor(cyl), data=mtcars, geom="bar") + coord_flip()
```

### Difference between fill/color bars
```{r fig.width=7, fig.height=4}
# difference between fill/color bars
qplot(factor(cyl), data=mtcars, geom="bar", fill=factor(cyl))
qplot(factor(cyl), data=mtcars, geom="bar", colour=factor(cyl))
```


```{r fig.width=7, fig.height=4}

# fill by variable
qplot(factor(cyl), data=mtcars, geom="bar", fill=factor(gear))
# use different display of bars (stacked, dodged, identity)
head(diamonds)
```

### Types of Bar Plots
```{r fig.width=7, fig.height=4}
qplot(clarity, data=diamonds, geom="bar", fill=cut, position="stack")
qplot(clarity, data=diamonds, geom="bar", fill=cut, position="dodge")
qplot(clarity, data=diamonds, geom="bar", fill=cut, position="fill")
qplot(clarity, data=diamonds, geom="bar", fill=cut, position="identity")
```

### Frequency Polygon

```{r fig.width=7, fig.height=4}
qplot(clarity, data=diamonds, geom="freqpoly", group=cut, colour=cut, position="identity")
```


```{r fig.width=7, fig.height=4}
qplot(clarity, data=diamonds, geom="freqpoly", group=cut, colour=cut, position="stack")
```

### Histograms

```{r fig.width=7, fig.height=4}
# histogram
qplot(carat, data=diamonds, geom="histogram")
```


```{r fig.width=7, fig.height=4}
# change binwidth
qplot(carat, data=diamonds, geom="histogram", binwidth=0.1)
qplot(carat, data=diamonds, geom="histogram", binwidth=0.01)
```
### Density Plot
```{r}
qplot(carat, data = diamonds, geom = "density", colour = color)
qplot(carat, data = diamonds, geom = "histogram", fill = color)
```

```{r fig.width=7, fig.height=4}
# use geom to combine plots (hint: order of layers)
qplot(wt, mpg, data=mtcars, geom=c("point", "smooth"))
qplot(wt, mpg, data=mtcars, geom=c("smooth", "point"))
qplot(wt, mpg, data=mtcars, color=factor(cyl), geom=c("point", "smooth"))
```


